---
title: "Introduction to cleanr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to cleanr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

devtools::load_all()
```

## Overview
The cleanr package provides a small set of helper functions for common data cleaning tasks. This vignette demonstrates a basic workflow using the functions currently available in the package

## Example data

```{r}
df <- data.frame(
  id = 1:10,
  height = c(120, 135, NA, 172, 180, 156, 192, 140, NA, 160),
  weight = c(58, 65, 73, 80, 78, 91, 96, 88, 85, NA),
  group = c("A", "B", "B", "A",  "A", "B", "B", "A", NA, "B")
)
df
```
## Functions

## Scaling numeric variables

The scale_numeric() function standardises numeric variables using z-scores

```{r}
scale_numeric(df)
```

If you do not wish to scale the id column, you can choose which columns you want to scale using cols

```{r}
scale_numeric(df, cols = c("height", "weight"))
```

## Detecting outliers
The detect_outliers() function identifies potential outliers in a numeric vector using the interquartile range rule

```{r}
detect_outliers(df$weight)
```

## Summary tables

The summary_table() function creates a summary table for variables in a data frame. For numeric variables it computes mean, sd, and quantiles. For categorical variables it reports counts and proportions by level

```{r}
summary_table(df)
```
## Removing missing values

The remove_missing() function removes observations with missing values from a data frame



